### 2024-05-30 Comparison.Rmd
reshaped plots

---
title: "Template Title"
author: "Niklas Hausmann"
date: "Today"
output: pdf_document
---
```{r Libraries,echo=FALSE,message = FALSE,warning = FALSE}

knitr::opts_chunk$set(echo=FALSE,message = FALSE,warning = FALSE)

{
  pacman::p_load(
    here,
    janitor,
    tidyverse,
    cowplot,ggx,ggpubr,
    patchwork,
    RColorBrewer)

  theme_set(theme_cowplot())
}

source("DTW_Script.R")

files <- list.files(path = here("data"), pattern = "\\.csv$", full.names = TRUE)

MgCa <- files %>%
  set_names() %>%
  map_dfr(~read_csv(.x, col_names = c("x", "y", "z", "mg_ca", "std", "relstd")), .id = "sampleID") %>%
  mutate(sampleID = tools::file_path_sans_ext(basename(sampleID))) %>%
  mutate(sampleID = str_replace(sampleID, "_analyzed_data", "")) %>%
  mutate(sampleID = str_replace(sampleID, "^(.*?A\\d+).*", "\\1")) %>% 
  mutate(sampleID = str_replace(sampleID, "-A10", "")) %>% 
  group_by(sampleID) %>% 
  mutate(dist=row_number()*0.01) %>% 
  ungroup() %>% 
  filter(!(sampleID=="ND-1016-4"&dist>6.5)) %>% 
  filter(relstd<0.1)

D18O <- read_csv(here("data","Isotopes","d18O.csv"))

```

# Plots

```{r}
#| label: fig-timeline
#| fig-cap: Timeline of recent earthquakes on La Palma
#| fig-alt: An event plot of the years of the last 8 eruptions on La Palma.
#| fig-height: 1.5
#| fig-width: 6

P1 <- MgCa %>% 
  filter(std<0.1) %>% 
  filter(mg_ca<0.7) %>% 
  filter(dist<10) %>% 
ggplot()+
  aes(dist,mg_ca)+
  geom_path(col="firebrick")+
  xlim(0,10)+
  facet_wrap(~sampleID,scales = "free_y",nrow=1)

P2 <- D18O %>% 
  filter(sampleID %in% MgCa$sampleID) %>% 
  ggplot()+
  aes(dist,d18O)+
  geom_path(col="cornflowerblue")+
  scale_y_reverse()+
  facet_wrap(~sampleID,scales = "free_y",nrow=1)

plot_compare <- P1/P2

```

# Compare

```{r}
D18O %>%
  filter(sampleID %in% MgCa$sampleID) %>%
  bind_rows(
    MgCa %>% filter(std < 0.1) %>%  filter(mg_ca < 0.7) %>%  filter(dist < 10)%>% select(sampleID, dist, mg_ca)
  ) %>%
  mutate(d18O = d18O * (-0.05)) %>%
  pivot_longer(values_to = "value",
               names_to = "proxy",
               cols = c(d18O, mg_ca)) %>%
  na.omit() %>%
  ggplot() +
  aes(dist, value, col = proxy) +
  geom_path() +
  facet_wrap( ~ sampleID, scales = "free")

```

```{r}
Samples <- unique(MgCa$sampleID)

# Initialize an empty list to store the results of shell_dtw for each sample
dtw_results <- list()

Corr_data <- tibble()

# for (sample in Samples) {
  temp_D18O <- D18O %>% 
    filter(sampleID == sample) %>% 
    select(d18O) %>% 
    rename(d18o = d18O) %>% 
    mutate(sample_seq = row_number())

  temp_MgCa <- MgCa %>% 
    filter(sampleID == sample) %>% 
    filter(!(sampleID=="ND-1016-4"&dist<0.3)) %>% 
    select(mg_ca) %>% 
    mutate(sample_seq = row_number())

  # Apply shell_dtw to the temp dataframes and store the result in the list
  dtw_results[[sample]] <- shell_dtw_data(temp_D18O, temp_MgCa,plot_alignment = FALSE)

   # Add the sampleID column to temp_data
  temp_data <- dtw_results[[sample]]$data %>% 
    mutate(sampleID = sample)

  # Append temp_data to Corr_data
  Corr_data <- bind_rows(Corr_data, temp_data)
# }

```

# Show Correlation data

```{r}

P_all <- Corr_data %>% 
  pivot_wider(values_from = "value",names_from = "proxy") %>% 
  ggplot()+
  aes(d18o,mg_ca,col=sampleID)+
  geom_point(show.legend = FALSE)+
  labs(x=expression("‰ δ" ^ "18" * "O [VPDB]"),y="Mg/Ca intensity ratio")+
  stat_smooth(aes(fill = sampleID,col=sampleID), method = "lm", formula = y~x,show.legend = FALSE) +
  # stat_regline_equation(label.x = 2, #label.y = 32
  #   aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
  #   formula = y~x
  # ) +

  scale_colour_manual(values = c("#007f7f", "#d73027", "#ffc100","black", "black"))+
  scale_fill_manual(values = c("#007f7f", "#d73027", "#ffc100","black", "black"))
# Expanded colour palette with 6 additional colours
# Updated colour palette with 4 new replacements for the last colours
# Refined colour palette with 3 additional colours
# c("#007f7f", "#d73027", "#ffc100", "black", "#1a9850", "#542788", "#3288bd", "#99d8c9")

P_individual <- P_all+
  facet_wrap(~sampleID,ncol=2)+#ylim(0.15,NA)+
  ggpubr::stat_regline_equation(label.x = 1.5, label.y = 0.25,
    aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),show.legend = FALSE,
    formula = y~x
  )

Plot_Correlations <- P_all/P_individual+plot_layout(heights = c(0.5,0.5))

# ggsave(plot=Plot_Correlations, here("Figures", "Correlations.png"), units = "cm", width = 29.7, height = 20.7)
```

# Look at ND-1016-4
```{r}

# sample <- "ND-1016-4"

temp_D18O <- D18O %>% 
    filter(sampleID == sample) %>% 
    rename(d18o = d18O) %>% 
    ggplot() +
  aes(dist, d18o) +
  scale_y_reverse()+
  geom_path(col="turquoise")

temp_MgCa <- MgCa %>% 
    filter(sampleID == sample) %>% 
    filter(dist>0.3) %>%
    ggplot()+
  aes(dist,mg_ca)+
  geom_path(col="coral")

temp_MgCa/temp_D18O
```