### 2024-06-27 Manuscript.tex
cleaned up format

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Palaeo3}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Interspecies comparisons of Mg/Ca ratios in limpet shells},
  pdfauthor={Niklas Hausmann; Donna Surge; Ivan Briz-Godino},
  pdfkeywords={Sclerochronology, Limpets, Elemental Ratio, Mg/Ca},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Interspecies comparisons of Mg/Ca ratios in limpet shells}
\author[1]{Niklas Hausmann%
\corref{cor1}%
}
 \ead{niklas@palaeo.eu} 
\author[2]{Donna Surge%
%
}

\author[3]{Ivan Briz-Godino%
%
}

\affiliation[1]{organization={Leibniz Zentrum für
Archäologie},addressline={Ludwig-Lindenschmit-Forum
1},city={Mainz},country={Germany},countrysep={,},postcode={55116},postcodesep={}}
\affiliation[2]{organization={University of North
Carolina},addressline={104 South Road, 225 Geology
Building},city={Chapel Hill,
NC},country={US},countrysep={,},postcode={27599-3315},postcodesep={}}
\affiliation[3]{organization={},country={Spain},countrysep={,},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This study provides a short reassessment of the use of Magnesium to
Calcium (Mg/Ca) ratios in Atlantic limpet shells to determine past sea
surface temperatures. While \emph{Patella vulgata} along the Spanish
shoreline has since then repeatedly produced reliable correlations
between sea surface temperature and Mg/Ca ratios, this is not the case
for other patelloid species. \emph{Patella vulgata} and \emph{Nacella
deaureata} have been studied using Mg/Ca with mixed or contrary results.
In this study, we present elemental maps of various such species
together with stable oxygen isotope values. Our dataset also includes
specimens that were previously unsuccessful in providing significant
correlations between \(\delta\)\textsuperscript{18}O and Mg/Ca ratios.
By reassessing these previous specimens and including a wider range of
modern and archaeological samples (n=12) from three patelloid species
(\emph{P. vulgata}, \emph{N. deaureata}, and \emph{N. magellanica}) we
further add to the growing set of evidence for the reliable use of Mg/Ca
ratios to detect palaeotemperature change and serve as a means to
determine ontogenetic age and season of capture as well as to reveal
locations of interest within the growth record (i.e.~annual temperature
minima and maxima) for the targeted analysis using
\(\delta\)\textsuperscript{18}O or clumped oxygen isotope analysis.
\end{abstract}

\begin{keyword}
    Sclerochronology \sep Limpets \sep Elemental Ratio \sep 
    Mg/Ca
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{Introduction}

Limpet shells are commonly found within archaeological sites and and
past shorelines due to their robust carbonate structure and the
long-term use of limpets as a marine food source. They have successfully
been studied in the past in the context of coastal subsistence
economies, site occupation on seasonal
\citep{Shackleton1973-ij, Parker2018-wf, Bosch2018-ud} or long-term
scale \citep{Ortiz2015-mr}, as well as palaeotemperature
\citep{Fenger2007-gf, Surge2012-ba, Wang2012-ee, Colonese2012-ct, Ferguson2011-zl}.
Determining past sea surface temperature (SST) change largely relies on
the measurement of \(\delta\)\textsuperscript{18}O-values within the
calcitic parts of the shell, but attempts have been made to also use
elemental ratios, such as magnesium to calcium (Mg/Ca), to have an
alternative measure, that potentially provides a more accurate SST
estimate than \(\delta\)\textsuperscript{18}O-values, which are also
affected by changes in salinity. In addition, the data acquisition of
elemental ratios --- either through
laser-ablation-isotope-ratio-mass-spectrometry (LA-ICP-MS) or laser
induced breakdown spectroscopy (LIBS), can be much faster and
cost-effective, increasing the number of specimens that can be studied
overall \citep{Durham2017-fh, Hausmann2023-ih}.

Problematically, links between Mg/Ca ratios and SST changes in many
marine mollusc shells, including limpets, have shown to be unreliable
\citep{Surge2008-ri, Wanamaker2008-zl, Schone2010-yl, Freitas2012-tx, Graniero2015-zv, Poulain2015-dg, Vihtakari2017-wd}.
This is particularly the case where there is little available additional
information on metabolic processes, organic components of the shell
matrix, intra-increment and intra-shell variability, and growth rates,
which can independently and unpredictably affect Mg/Ca ratios and
confound their interpretation as temperature proxy. Confusingly,
multiple different temperature equations have been found for the same
species (see \citep{Freitas2012-tx, Vihtakari2017-wd} and references
therein). Coeval specimens sharing one locality can also show
differences in their relation to SST \citep{Hausmann2019-fi}. Where the
use of Mg/Ca as a palaeotemperature proxy was successful, anomalous
patterns in some specimens had still to be filtered out by hand,
reducing the overall robustness of the results of those successful
studies \citep{Ferguson2011-zl}.

While recent research particularly of \emph{Patella} sp. in the
Mediterranean and Southwest Europe have provided promising results
\citep{Hausmann2019-fi, Garcia-Escarzaga2015-jc, Garcia-Escarzaga2018-nf}.
There remains a lack of clarity for Atlantic limpet species,
particularly since it was last shown here, that they are not reliable
recorders of palaeotemperature \citep{Graniero2015-zv}.

In this study we will repeat and expand the analysis of Atlantic limpets
to determine the reliability of Mg/Ca ratios as palaeotemperature
proxies. To do this, we sampled a set of previously published and
unpublished limpet specimens dating to modern and archaeological
contexts using LIBS. LIBS allows us to carry out 2D imaging of entire
shell sections, which helps us to navigate the complex elemental
structure of the shell and better separate the external,
temperature-related changes from the internal and less understood
factors that influence the Mg/Ca ratio.

By relying and adding onto published datasets, we were able to
simultaneously avoid costs for new high-resolution
\(\delta\)\textsuperscript{18}O-data, to use real-world examples, and to
also provide pilot-data for areas of existing research interest.
Generally establishing the usefulness and reliability of Mg/Ca as SST
proxy in limpets should help to provide a platform for future research
and an important stepping stone to better understand elemental ratios in
other marine mollusc shells.

\section{Materials and Methods}\label{Methods}

\subsection{Limpet specimens}\label{limpet-specimens}

The analysed specimens, their origin and respective studies with
research background information can be found in
Table\hyperref[Table_1]{1}. Here we will briefly summarise their
contextual information, which can be accessed in more detail at the
respective studies
\citep{Nicastro2020-ih, Surge2012-ba, Graniero2017-io}. While those
studies also included other specimens, their accessibility or state of
preservation did not all lend them to be re-analysed.

\phantomsection\label{Table_1}
\fontsize{8pt}{8pt}\selectfont

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1224}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2449}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1633}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1224}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2041}}@{}}
\caption{Overview of the modern and archaeological limpet specimens
analysed in this study.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Modern & \citep{Nicastro2020-ih} & \emph{N. deaureata} & Cambaceres Bay
& ND-1016-3 & Stable oxygen \\
& & & - Tierra del & ND-1016-4 & and carbon \\
& & \emph{N. magellanica} & Fuego (AR) & NM-1016-1 & isotope analysis \\
& & & & NM-1016-3 & \\
& & & & & \\
& \citep{Graniero2017-io} & \emph{P. vulgata} & Rack Wick Bay - &
ORK-LT5 & Stable oxygen \\
& & & Westray (UK) & & and carbon \\
& & & & & isotope analysis; \\
& & & & & Mg, Li, Sr, Ca. \\
& & & & & \\
Archaeological & \citep{Surge2012-ba} & \emph{P. vulgata} & Rack Wick
Bay - & QG1-7188-1 & Stable oxygen \\
& & & Westray (UK) & QG1-7189-2 & and carbon \\
& & & & QG2-1061-1 & isotope analysis \\
& & & & QG2-1064-1 & \\
& & & & QG2-7180-1 & \\
& & & & QG2-7180-2 & \\
& & & & QG1-7246-1 & \\
\end{longtable}

\normalsize

There are two main locations that we sourced specimens from: the Beagle
Channel in Tierra del Fuego (Argentina) and the island of West Ray in
Orkney (UK). Four specimens were collected from the Beagle Channel in
Cambaceres Bay in October 2016. The area receives around 570 mm
precipitation but the channel waters are predominantly influenced by
marine currents and in the period of October 2015 to October 2016 stayed
around 30.7 ± 0.7 psu \citep{Nicastro2020-ih}. SSTs range from 4.7ºC in
August to 10ºC in February.

On Westray one modern shell (ORK-LT5) was collected in Rack Wick Bay in
August 2009. Westray lies about 70 km north of the Scottish mainland and
experiences virtually no freshwater input with salinity values of 34.0
-- 34.5 psu \citep{Inall2009-ho}. The annual SST ranges from 6.3ºC in
March to 13.8ºC in August (Figure~\ref{fig-SSTs}).

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-SSTs-1.pdf}

}

\caption{\label{fig-SSTs}Monthly Sea Surface Temperature values averaged
from years. SST values were sourced from \citet{Good2020-nl}.}

\end{figure}%

In total, we studied 7 archaeological specimens of \emph{P. vulgata}
coming from Quoygrew on Westray dating to 900 -- 1200 CE. These shells
have previously been studied to characterise seasonal temperature change
during the Medieval Climate Anomaly and the Little Ice Age
\citep{Surge2012-ba}. Together they show a range of lifespans, with some
comparatively long for their size (QG1-7246-1, 12 years within 16 mm of
growth) making them particularly susceptible to time-averaging. That
said, the majority are under 5 years old. These shells are an ideal
comparative dataset from an archaeological assemblage to compare to the
modern shell ORK\_LT5 \citep{Graniero2017-io}.

\subsection{Methods}\label{methods}

\subsubsection{Mg/Ca ratios}\label{mgca-ratios}

Mg/Ca elemental imaging was carried out using Laser Induced Breakdown
Spectroscopy at the Leibniz Zentrum für Archäologie (Mainz-Germany)
using previously published methods \citep{Hausmann2023-ih}. These
involve the ablation of the shell section using an infrared (1064 nm)
laser (1--2 mJ, 100 Hz) onto a surface area of 20--30 µm to generate a
plasma plume. This plume emits light, which is measured using a
synchronised spectrometer. The resulting light spectrum quantifies the
emission lines of magnesium (MgII; 279.553 nm) and calcium (CaII;
315.887 nm) to determine their intensity ratio. While these two peaks
alone do not represent the molar concentrations of both elements (as
opposed to calibration free LIBS; e.g.
\citep{Martinez-Minchero2022-jz}), the intensity ratio is linearly
correlated to the molar concentration \citep{Hausmann2017-oa} and can be
used as a reliable indicator of Mg/Ca changes within the shell
carbonate.

Using this system we carried out elemental imaging of the shell
specimens at a resolution of 50 µm distance between sampling spots. Each
spot was irradiated 10 times with the 3 first spectra used for cleaning
steps and the remaining summed to get an average Mg/Ca intensity ratio
for each sample spot. Subsequently, we re-sampled the section using a
line scan at 10 µm resolution. This leads to an overlap between sample
spots, but also allows for a continuous record without gaps. Intensity
ratios were filtered for cases with high relative standard deviation
(i.e.~more than 10\%). These occurred in places where the previous
sampling procedures for carbonate powder as part of the oxygen isotope
analysis left an uneven sample surfaces introducing variability in the
plasma generation and thus uncertainty in the data of these locations.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_iso-1.png}

}

\caption{\label{fig-Nac_iso}Isotope values of \emph{Nacella} sp. limpet
shells from Cambaceres Bay.}

\end{figure}%

\subsubsection{Oxygen isotopes}\label{oxygen-isotopes}

High-quality oxygen isotope values (\(\delta\)\textsuperscript{18}O)
were acquired from previous publications on \emph{P. vulgata}
\citep{Surge2012-ba, Graniero2017-io}, \emph{N. deaureata} and \emph{N.
magellanica} \citep{Nicastro2020-ih}, in which detailed descriptions of
the stable isotope analysis methodology can be found. In short, all
shells were sectioned along the main growth axis to expose the internal
growth structure. From the exposed section carbonate powder samples were
acquired using a micromill with resolutions of 100--200 µm and by
targeting the calcitic M+2 layer of the shells. Analytical precision of
the isotope values was consistently at 0.05--0.10‰ and values are
reported in per mil units (‰) relative to the VPDB (Vienna Pee Dee
Belemnite) standard.

Figure~\ref{fig-Nac_iso} and Figure~\ref{fig-Pat_iso} show the
previously acquired sequential \(\delta\)\textsuperscript{18}O-values of
limpet shells by species. Due to the high sampling resolution, all
shells show seasonally resolved and multi-year quasi-sinusoidal records
of SST change , with \emph{Nacella} sp. shells living 2--3 years and
\emph{Patella v}. regularly over 3 years. In those instances where
longer lifetimes are encountered (i.e.~specimen QG1-7246-1,
Figure~\ref{fig-Pat_iso} b).

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_iso-1.png}

}

\caption{\label{fig-Pat_iso}Isotope values of \emph{Patella vulgata}
limpet shells from Quoygrew (QG, \emph{a-g}) and Rack Wick Bay (ORK,
\emph{h}) on Westray, Orkney.}

\end{figure}%

Interestingly, the annual ranges in
\(\delta\)\textsuperscript{18}O-values are not similar between the
coeval \emph{Nacella} sp. specimens with \(\delta\)\textsuperscript{18}O
values for the previous summer temperatures ranging from 0.4‰ to above
1.0‰, and for winter values between 2.7‰ and 2.0‰. \emph{P. vulgata} has
similarly variable \(\delta\)\textsuperscript{18}O-values for summer and
winter seasons, alas the lack of more that one modern specimens, means
that we cannot quantify the variability of recorded annual ranges
between specimens.

\subsection{Dynamic Time Warping}\label{dynamic-time-warping}

Because the exact location used in the analysis of stable oxygen
isotopes is not possible to reconstruct for each carbonate powder
sample, we used dynamic time warping (DTW) to align the time series of
\(\delta\)\textsuperscript{18}O-values and Mg/Ca ratios. DTW is an
algorithm that measures similarity between two proxy-sequences, which
may vary in sampling resolution or interval. By stretching or
compressing sections of the series, DTW finds the probable alignment
between the two sequences. This allows us to compare the proxy data sets
more effectively, ensuring that the temporal dynamics of each shell are
accurately matched despite possible discrepancies in sampling intervals
or rates. We applied the DTW algorithm using the \texttt{dtw} package in
R \citep{Giorgino2009-sj, R_Core_Team2020-mk}, which provides a robust
framework for aligning time series data. This involved selecting
appropriate distance measures and constraints to ensure meaningful
alignment. The process involved iterative adjustments to minimise the
overall distance between corresponding points in the two data sets.

\section{Results}\label{Results}

\subsection{\texorpdfstring{\emph{Patella
vulgata}}{Patella vulgata}}\label{patella-vulgata}

The elemental imaging of \emph{P. vulgata} shells consistently showed
repeating patterns of Mg/Ca intensity ratio changes with high Mg/Ca
intensity ratios indicating and low ratios indicating low temperature
periods. As examples, Figure~\ref{fig-Pat_LIBS} shows the
2D-distribution of elemental ratios across the entire section of ORK-LT5
and the preserved anterior side of QG1-7188-1 (additional maps can be
found in the supplementaries). The repeating patterns are found across
the calcitic layers that are exterior to the myostracum (layers m+2 and
m+3) as well as the layers that are interior (m-2)
\citep{Fenger2007-gf}. The layer m-1, which is also interior, consists
of aragonite and is seen in Figure~\ref{fig-Pat_LIBS} in grey, as its
Mg/Ca intensity ratio consistently falls below the range of interest
(0.2 and above). Compared to other patelloid shells, whose interior is
almost entirely made of aragonite, this layer is very thin.

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_LIBS-1.pdf}

}

\caption{\label{fig-Pat_LIBS}Example LIBS maps and line scans of
analysed specimens. (\emph{a+b}) ORK-LT and (\emph{c}+\emph{d})
QG1-7188-1. Note that the ranges of Mg/Ca intensity ratios are not
identical but are chosen arbitrarily to increase contrast. Areas in grey
consist of low-Mg aragonite and fall outside the chosen colour-range.}

\end{figure}%

Both specimens in Figure~\ref{fig-Pat_LIBS} show increased Mg/Ca
intensity ratios towards the outside of the shell and some degree of
intra-increment variability. This is particularly visible in
Figure~\ref{fig-Pat_LIBS} c, where Mg/Ca intensity ratios range from
\textasciitilde0.5--0.7 in the first year of growth (9--12 mm on the
y-axis). A similar pattern is visible in the interior m-2 layer of
ORK-LT5 (Figure~\ref{fig-Pat_LIBS} a), which increase towards the
anterior of the shell (0--10 mm on the x-axis).

Line scans also indicate well the quasi-sinusoidal change of Mg/Ca
intensity ratios expected based on the stable isotope data and
elemental-imaging. That said, some of the variability of Mg/Ca intensity
ratios within one season such as the summer period of QG1-7188-1 between
0.5 and 2.0 mm distance to the shell edge (Figure~\ref{fig-Pat_LIBS} d),
is visible in a more dramatic manner than it appears in the 2D-image,
showing the downside of line-scanning as opposed to the broader milling
approach used for the analysis of carbonate powder
\citep{Ferguson2011-zl}.

Figure~\ref{fig-Pat_Comp} shows the line scan of ORK-LT5 in comparison
to its \(\delta\)\textsuperscript{18}O-values (similar graphs for other
specimen can be found in the supplementaries). The measurements of the
distance to the shell edge are not entirely identical, most likely
because one growth increment can have several different distances.
Depending on where one measures the distance and a line scan towards the
interior would be shorter than a line scan towards the exterior. That
said, increases and decreases of both proxies seem to mirror each other
and to be well aligned.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_Comp-1.png}

}

\caption{\label{fig-Pat_Comp}Comparison of LIBS line-scan and sequential
\(\delta\)\textsuperscript{18}O-values of specimen ORK-LT5}

\end{figure}%

Aligning the records programmatically using dynamic time warping,
allowed us to compare them directly to determine specimen-specific
equations and quantify the correlation of both proxies
(Figure~\ref{fig-Pat_Corr}). The equations for all specimens are
different but the majority (barring one: QG1-7246-1) do seem to cluster
around shared parameters (\(\delta\)\textsuperscript{18}Ο = 0.6 -- 0.065
* Mg/Ca). The various coefficients of determination
(R\textsuperscript{2}) range between 0.81 (QG1-7246-1) and 0.95
(QG1-7189-2). These specimens are also the oldest and youngest,
respectively, specimens, suggesting that lower growth rates and time
averaging had a negative effect on the correlation of both proxies. The
mean R\textsuperscript{2} value is 0.89, suggesting a generally good fit
between the two proxies.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_Corr-1.png}

}

\caption{\label{fig-Pat_Corr}Correlation graphs for \emph{Patella
vulgata} specimens.}

\end{figure}%

\subsection{\texorpdfstring{\emph{Nacella}
sp.}{Nacella sp.}}\label{nacella-sp.}

The LIBS data for \emph{Nacella} sp. specimens is less straightforward
with shells being much thinner and thus patterns being more difficult to
make out. Both specimens (ND-1016-3 and NM-1016-3) shown as example in
Figure~\ref{fig-Nac_LIBS} (see additional graphs in supplementaries)
also experience increases of Mg/Ca intensity ratio towards the exterior
of the shell, similar to the \emph{P. vulgata} specimens from Orkney.
This intra-increment heterogeneity combined with the thinness of the
shell, further complicates their analysis compared to \emph{P. vulgata}
or other \emph{Patella} species \citep{Hausmann2019-fi}. Nevertheless,
repeating patterns were visible in the anterior section of the shell,
which were also mirrored in the inner layers (m-2) of specimens
NM-1016-3 (Figure~\ref{fig-Nac_LIBS} c).

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_LIBS-1.pdf}

}

\caption{\label{fig-Nac_LIBS}Example LIBS maps and line scans of
analysed specimens. (\emph{a+b}) ND-1016-3 and (\emph{c}+\emph{d})
NM-1016-3. Note that the ranges of Mg/Ca intensity ratios are not
identical but are chosen arbitrarily to increase contrast. Areas in grey
consist of low-Mg aragonite and fall outside the chosen colour-range.
Note the inset in (\emph{c}) where the shell edge has been resampled
using a higher resolution (50 µm instead of 100 µm) to better understand
the Mg/Ca ratios at the shell edge.}

\end{figure}%

Interestingly, the Mg/Ca intensity ratios in the \emph{Nacella} sp.
shells were much lower (0.05--0.30) than those seen in other patelloid
species (e.g.~0.3--0.7 in \emph{Patella vulgata} above, or 0.5--1.5 in )
using the same emission lines. This is particularly visible in
Figure~\ref{fig-Nac_LIBS} c, where Mg/Ca intensity ratios range from
\textasciitilde0.5--0.7 in \citep{Hausmann2023-ih}). While the intensity
ratio depends chiefly on the chosen emission lines to calculate the
proportion of the chosen Magnesium to the Calcium peak, the low ratios
still indicate lower concentrations than seen elsewhere, or seen only in
aragonitic parts of e.g.~\emph{Patella} shells. Similar observations
have been made before \citep{Graniero2017-io}.

Line scans reflect well the changes previously indicated by
\(\delta\)\textsuperscript{18}O values and also indicate about two years
of recorded growth. That said, the 2D imaging suggests that in both
specimens a third summer can be added to the total record, which is not
captured by the line scans. The variability seen in the \emph{P.
vulgata} specimens above is not as strong, which might also be related
to the fast and consistent growth rate of the shells.

Figure~\ref{fig-Nac_Comp} shows the line scans of ND-1016-3 and
NM-1016-3 in comparison to their \(\delta\)\textsuperscript{18}O-values
(similar graphs for the other \emph{Nacella} sp. specimen can be found
in the supplementaries). Compared to the \emph{P. vulgata} records,
these mirror each other much more clearly, due to the brief period
recorded in the shells and its simple make-up.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_Comp-1.png}

}

\caption{\label{fig-Nac_Comp}Comparison of LIBS line-scan and sequential
\(\delta\)\textsuperscript{18}O-values of specimen ORK-LT5}

\end{figure}%

The \emph{Nacella} sp. specimens showed similarly clustered
correlations, which interestingly were not grouped by species. In fact,
specimens ND-1016-3 (\emph{Nacella deaureata)} and NM-1016-3
(\emph{Nacella magellanica}) are almost identical in their fitted
equations (Figure~\ref{fig-Nac_Corr}). Both also share a high
coefficient of determination (R\textsuperscript{2}) of 0.95 and 0.92
respectively, with the remaining specimens being somewhat lower
(R\textsuperscript{2} = 0.76 for ND-1016-4 and R\textsuperscript{2}=0.84
for NM-1016-3). The mean R\textsuperscript{2} value for each
\emph{Nacella} species is 0.86 for \emph{Nacella deaureata} and 0.88 for
\emph{Nacella magellanica}. Why two specimens of different species are
so well aligned is not entirely certain and randomness cannot for now be
ruled out. That said, since they both have a high R\textsuperscript{2}
value, it might just be that they were both minimally affected by
factors other than SST, including the sampling location of our line
scans.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_Corr-1.png}

}

\caption{\label{fig-Nac_Corr}Correlation graphs for \emph{Nacella} sp.
specimens}

\end{figure}%

\section{Discussion}\label{Discussion}

The Mg/Ca ratios from the 2D elemental imaging present positive results
on SST being the main factor behind Mg/Ca concentration in three
Atlantic limpet species. Annually repeating patterns of increasing
(spring/summer) and decreasing (autum/winter) SST were also visible in
the line scans, albeit less straightforward than in the 2D-data.
Nevertheless, the resulting R\textsuperscript{2}-values for each species
were high (\emph{P. vulgata mean R\textsuperscript{2}}=0.89, \emph{N.
deaureata mean R\textsuperscript{2}}=0.86\emph{, N. magellanica mean
R\textsuperscript{2}}=0.88). These results are similar to previously
published data on \emph{Patella} sp. shells (Table
\hyperref[Table_2]{2}). R\textsuperscript{2} values of \emph{P.
depressa} lie between 0.78 and 0.87 \citep{Garcia-Escarzaga2021-ij}, for
\emph{P. caerulea} from the Mediterranean between 0.78 and 0.96 ---
although with one outlier of 0.33 (specimen MP64A), which had increased
Mg/Ca ratios in increments with rich organic contents
\citep{Hausmann2019-fi}.

\phantomsection\label{Table_2}
\fontsize{8pt}{8pt}\selectfont

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2083}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3472}}@{}}
\caption{Table showing correlations of Mg/Ca and
\(\delta\)\textsuperscript{18}O found in other studies. * indicates that
only SST data and no other geochemical data was used. ** these samples
were used to determine a shared R\textsuperscript{2}-value of
0.79.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Locality
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Specimen
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
R\textsuperscript{2}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Locality
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Specimen
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
R\textsuperscript{2}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\emph{P. depressa} & Northern Spain & LAN541 & 0.87 &
\citep{Garcia-Escarzaga2021-ij} \\
& & LAN545 & 0.86 & \\
& & LAN554 & 0.78 & \\
& & LAN559 & 0.82 & \\
\emph{P. caerulea} & Croatia & ISTPC1 & 0.9 & \citep{Hausmann2019-fi} \\
& & ISTPC2 & 0.84 & \\
& Crete & AF1911A & 0.91 & \\
& & AF3003A & 0.92 & \\
& Israel & AKKPC2 & 0.96 & \\
& & AKKPC3 & 0.89 & \\
& & FRMPC1 & 0.84 & \\
& & FRMPC2 & 0.96 & \\
& Libya & MO31A & 0.83 & \\
& & MP64A & 0.33 & \\
& & MP67A & 0.96 & \\
& & MP68A & 0.81 & \\
& Malta & MA10 & 0.82 & \\
& Tunisia & TUNPC1 & 0.81 & \\
& & TUNPC2 & 0.78 & \\
& Turkey & ANTPC1 & 0.95 & \\
& & ANTPC2 & 0.93 & \\
& & KIZPC1 & 0.94 & \\
& & KIZPC2 & 0.86 & \\
\emph{P. rustica} & Gibraltar & JL1 & 0.02 & \citep{Ferguson2011-zl} \\
& & JL2 & 0.8** & \\
\emph{P. caerulea} & Gibraltar & JM00 & 0.69** & \\
& & JM30 & 0.83** & \\
\emph{P. vulgata} & Orkney & ORK-LT5 & not reported, here 0.88 &
\citep{Graniero2017-io} and this study \\
\end{longtable}

\normalsize

\emph{P. rustica} has two studied specimens (JL1 and JL2) with
R\textsuperscript{2}-values of 0.02 and 0.8, respectively
\citep{Ferguson2011-zl} (Figure~\ref{fig-Ferg}). Together with the data
from two \emph{P. caerulea} shells (JM00 and JM30 with
R\textsuperscript{2}-values of 0.69 and 0.83), the data from JL2 was
used determine a general Mg/Ca-SST relationship with an
R\textsuperscript{2}-value of 0.79. JL1 was disregarded in that
calculation, as well as the second year of growth in JM00, which
produced anomalous data. While no additional information about those
specimens is available, it seems likely that the Mg/Ca ratios from JM00,
which were interpreted as an ontogenetic trend, were potentially
influenced by the same intra-increment heterogeneity we see in some of
our shell specimens as well. Since these shells were analysed by
homogenising the carbonate powder of the m+2 layer and the more exterior
m+3 layer, these heterogeneities could potentially increased the Mg/Ca
ratio unpredictably, leading to what might look like an ontogenetic
trend in a linear dataset. Whether these anomalies affected JL1 in the
same way, albeit stronger, is not clear but it might well be possible.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Ferg-1.png}

}

\caption{\label{fig-Ferg}Correlation graphs for Ferguson et al., 2011
specimens.}

\end{figure}%

Another shell whose previous results we aim to explain further is
ORK-LT5, which has previously been shown to have no substantial
correlation with \(\delta\)\textsuperscript{18}O or SST in implication
\citep{Graniero2017-io}. Different to other specimens of this study, we
were able to resample this specimen and getting Mg/Ca data on a slightly
higher resolution (20--30 µm spots in a 100 µm raster compared to 50 µm
spots spaced at 150--300 µm) and using 2 dimensions, as well as using a
continuous high-resolution line-scan at (20--30 µm spots spaced at 10
µm) (Figure~\ref{fig-ORK_sub}). These additional data revealed
previously missed summer peaks at around 2--3 mm distance to the shell
edge (grey box in Figure~\ref{fig-ORK_sub}) and revealed additional
growth at the very edge (dashed box in Figure~\ref{fig-ORK_sub})
mirroring the change towards warmer temperatures at the time of
collection (August). Including these missing parts lead to a better
alignment of both records and a resulting R\textsuperscript{2}-value of
0.88.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-ORK_sub-1.pdf}

}

\caption{\label{fig-ORK_sub}Comparison of previous and new Mg/Ca data
for ORK-LT5 explaining the previously unsuccessful correlation of Mg/Ca
and \(\delta\)\textsuperscript{18}O-values}

\end{figure}%

While we were not able to resample the \emph{Nacella} sp. specimens in
Graniero \emph{et al.} \citeyearpar{Graniero2017-io}, we were able to
source other modern and archaeological specimens of the same species
from Tierra del Fuego, which can serve as an indicator. Similar to other
patelloid shells, the results of the \emph{Nacella} specimens in our
study were successful with high R\textsuperscript{2}-values (0.76--0.95)
and suggest Mg/Ca as a reliable proxy for SST (Figure~\ref{fig-Nac_Comp}
and supplementaries). Why these results are different from the studied
\emph{Nacella} sp. specimens in Graniero et al.
\citeyearpar{Graniero2017-io} is not obvious, but the occasional
alignment of the Mg/Ca and \(\delta\)\textsuperscript{18}O records of
their shells (Figure 6 in \citep{Graniero2017-io} ) indicate that there
is some relationship that --- potentially due to the low sampling
resolution or intra-increment heterogeneity --- fell apart and lead to
an overall result of an uncorrelated relationship between Mg/Ca and SST.

\section{Conclusion}\label{conclusion}

This study presents high-resolution and 2D information on Mg/Ca
concentrations in modern and archaeological patelloid shells and their
relation to sea surface temperature. The results for \emph{Patella
vulgata}, \emph{Nacella deaureata}, and \emph{Nacella magellanica} were
consistently promising with mean R\textsuperscript{2}-values of 0.89,
0.86, and 0.88 respectively.

LIBS elemental imaging provided some insight into elemental anomalies
found particulary in the external parts (m+3 layer) of the shell, which
introduce intra-incremental heterogeneity and which cannot be reconciled
with modern SST changes or SST changes recorded in the
\(\delta\)\textsuperscript{18}O of the same shells. Our 2D approach
allowed us to avoid these areas of intra-increment heterogeneity and
focus on areas that are predominantly controlled by changes in SST.

These results are particularly promising for future analyses of
patelloid shells for the study of season of collection in archaeological
sites as well as the reconstruction of sub-annual climatic conditions in
the past. Shells of the species studied here are widely available along
the European Atlantic façade and the Argentinian coastline and make up a
large part of marine mollusc remains in archaeological sites of those
regions. Improving their analysis in terms of cost and time efficiency
through the benefits of LIBS, which can be used in this context as a
screening tool for targeted \(\delta\)\textsuperscript{18}O analysis
focusing on annual minima and maxima revealed in the Mg/Ca record,
allows researchers to increase sample numbers and produce larger and
more in-depth climatic studies than previously possible.

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}