### 2024-06-28 Extract data.Rmd
edits

---
title: "Template Title"
author: "Niklas Hausmann"
date: "Today"
output: pdf_document
---

# Libraries 
```{r Libraries,echo=FALSE,message = FALSE,warning = FALSE}
knitr::opts_chunk$set(echo=FALSE,message = FALSE,warning = FALSE)

{
  pacman::p_load(
    here,
    janitor,
    tidyverse,
    cowplot,ggx,
    patchwork,
    RColorBrewer)

  theme_set(theme_cowplot())
}
```

# Data

```{r}
MgCa <- read_csv(here("data","MgCa.csv"))

NM_1016_1 <- MgCa %>% 
  filter(sampleID=="NM-1016-1") #%>% 
  # select(-sampleID,-dist)

NM_1016_1 %>% 
  filter(std<0.01) %>% 
  ggplot()+
  aes(x=dist,y=mg_ca)+
  geom_path()

# write_csv(QG1_7246,file = here("data","export","QG1_7246.csv"))

```