### 2024-09-13 Supplementary files 2 - isotope graphs.qmd
Corrections

---
title: "Comparison of geochemical datasets for “*Confirmation of Mg/Ca ratios as palaeothermometers in Atlantic limpet shells*”"
format: docx
---

# Overview
These supplementary files include the linescan information for Mg/Ca and $\delta$^18^O records of all analysed [*Patella vulgata*](#Patella) and [*Nacella* sp.](#Nacella) shells, which were analysed to reassess the use of Magnesium to Calcium (Mg/Ca) ratios as palaeothermometers in Atlantic limpet shells. These supplementary files are organised by species and specimen identifiers.

We also show correlation graphs of

```{r Library and datasets}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE

knitr::opts_chunk$set(echo=FALSE,message = FALSE,warning = FALSE)
# Packages
  pacman::p_load(
    here,
    janitor,
    tidyverse,
    cowplot,ggx,ggpubr,
    patchwork,
    RColorBrewer)

  theme_set(theme_cowplot(10))

# Scripts
# source("DTW_Script.R")
source("DTW_Script-update.R")

# Files

D18O <- read_csv(here("data","D18O.csv")) %>% 
  mutate(period=case_when(startsWith(sampleID,"Q") ~ "Archaeological",TRUE ~ "Modern")) %>% 
  mutate(species = case_when(
      startsWith(sampleID, "ND") ~ "Nacella deaureata",
      startsWith(sampleID, "NM") ~ "Nacella magellanica",
      TRUE ~ "Patella vulgata"
    ))

MgCa <- read_csv(here("data","MgCa.csv"))

Ferguson_data <- read_csv(here("data","Ferguson-data.csv")) %>% 
  na.omit()

Corr_data <- read.csv(here("data","Corr_data.csv"))

#This below caused problems
#%>% 
  #   pivot_wider(values_from = "value",names_from = "proxy") %>%
  # left_join(D18O %>% select(sampleID) %>% unique(),by="sampleID") %>%
  # pivot_longer(cols=c(d18o,mg_ca),values_to = "value",names_to = "proxy") %>% 
  # arrange(x)

# # Check overlaps
# # Find common sample IDs
# intersect(MgCa$sampleID, D18O$sampleID)
# 
# # Find sample IDs only in MgCa
# setdiff(MgCa$sampleID, D18O$sampleID)
# 
# # Find sample IDs only in D18O
# setdiff(D18O$sampleID, MgCa$sampleID)

D18O <-   D18O %>% filter(sampleID %in% MgCa$sampleID)

```

## Geochemical records of shells in this study
```{r}
#| fig-height: 7
#| fig-width: 16

# Calculate the 2.5th and 97.5th percentiles for mg_ca within each sampleID
percentiles <- MgCa %>%
  group_by(sampleID) %>%
  summarise(lower = quantile(mg_ca, 0.025, na.rm = TRUE),
            upper = quantile(mg_ca, 0.975, na.rm = TRUE))

# Function to create the plots
create_plots <- function(sample) {
  lower <- percentiles %>% filter(sampleID == sample) %>% pull(lower)
  upper <- percentiles %>% filter(sampleID == sample) %>% pull(upper)

  P1.1 <-
    MgCa %>%
    filter(sampleID == sample) %>%
    filter(relstd < 0.1) %>%
    filter(mg_ca >= lower & mg_ca <= upper) %>%
    ggplot() +
    aes(dist, mg_ca) +
    geom_path() +
    theme_cowplot(17) +
    labs(x = "", y = "Mg/Ca intensity ratio", title = sample)

  P1.2 <-
    D18O %>%
    filter(sampleID == sample) %>%
    ggplot() +
    aes(dist, d18O) +
    geom_path() +
    theme_cowplot(18) +
    scale_y_reverse() +
    labs(x = "Distance to edge (mm)", y = expression(paste(delta^{18}, "O (\u2030)")))

   P2.1 <-
    Corr_data %>%
    filter(sampleID == sample) %>%
    filter(proxy == "mg_ca") %>%
    ggplot() +
    aes(x, value) +
    # geom_errorbar(aes(ymin = value - sd, ymax = value + sd, width = 1)) +
    geom_ribbon(aes(ymin = value - sd, ymax = value + sd), alpha = 0.5) +
    geom_path() +
    theme_cowplot(17) +
    labs(x = "", y = "")

  P2.2 <-
    Corr_data %>%
    filter(sampleID == sample) %>%
    filter(proxy == "d18o") %>%
          group_by(x) %>% 
      mutate(sd=sd(value)) %>% 
      ungroup() %>% 
    ggplot() +
    aes(x, value) +
    # geom_errorbar(aes(ymin = value - sd, ymax = value + sd, width = 0.2)) +
    geom_ribbon(aes(ymin = value - sd, ymax = value + sd), alpha = 0.5) +
    geom_path() +
    theme_cowplot(18) +
    scale_y_reverse() +
    labs(x = "Warped time index", y = "")

    P1.1 / P1.2 |  P2.1 / P2.2 + plot_annotation(tag_levels = "a")

}

# Loop through each sampleID and create the plots
unique_sampleIDs <- unique(MgCa$sampleID)

for(sample in unique_sampleIDs) {
  plot_combined <- create_plots(sample)
  print(plot_combined)
}

```

## Geochemical records of modern shells in other studies mentioned within the main text

``` {r Ferguson Data, dev='png'}
#| fig-cap: Correlation graphs for Ferguson et al., (2011) specimens.
#| fig-height: 5
#| fig-width: 16
#| label: fig-Ferg

Ferguson_data <- read_csv(here("data","Ferguson-data.csv")) %>% 
  na.omit()

Ferg_all <-
  Ferguson_data %>% 
  ggplot()+
  aes(d18o,mg_ca,col=shell)+
  geom_point(show.legend = FALSE)+
  labs(x=expression(paste(delta^{18},"O (\u2030)")),y="Mg/Ca (mmol/mol)")+
  stat_smooth(aes(fill = shell,col=shell), method = "lm", formula = y~x,show.legend = FALSE) +
  # stat_regline_equation(label.x = 2, #label.y = 32
  #   aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),
  #   formula = y~x
  # ) +
  scale_colour_manual(values = c(  "#EC7063", "#007f7f", "#1f78b4", "#542788","black", "#d73027", "#EC7063")) +
scale_fill_manual(values = c(  "#EC7063", "#007f7f", "#1f78b4", "#542788","black", "#d73027", "#EC7063"))

Ferg_individual <- Ferg_all+
  facet_wrap(~shell,ncol=4)+#ylim(0.15,NA)+
  ggpubr::stat_regline_equation(label.x = 0.3, label.y = 26,
    aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~~")),show.legend = FALSE,size=6,
    formula = y~x
  )+  
  theme_cowplot(20)+
  coord_fixed(ratio =0.22) +
  theme(
    strip.background = element_rect(fill = "white", colour = NA),
    strip.text = element_text(colour = "black"),
    panel.border = element_rect(colour="black")
  )
Ferg_individual
#Ferg_all/Ferg_individual+plot_layout(heights = c(0.5,0.5))

```

## References
* Ferguson, J.E., Henderson, G.M., Fa, D.A., Finlayson, J.C., Charnley, N.R., 2011. Increased seasonality in the western
mediterranean during the last glacial from limpet shell geochemistry. *Earth Planet. Sci. Lett.* 308, 325–333. doi:10.1016/j.
epsl.2011.05.054.