### 2024-10-14 Manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{lineno}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Palaeo3}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Confirmation of Mg/Ca ratios as palaeothermometers in Atlantic limpet shells},
  pdfauthor={Niklas Hausmann; Donna Surge; Ivan Briz i Godino},
  pdfkeywords={Sclerochronology, Patella, Nacella, LIBS, Mg/Ca},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Confirmation of Mg/Ca ratios as palaeothermometers in Atlantic
limpet shells}
\author[1]{Niklas Hausmann%
\corref{cor1}%
}
 \ead{niklas@palaeo.eu} 
\author[2]{Donna Surge%
%
}

\author[3]{Ivan Briz i Godino%
%
}

\affiliation[1]{organization={Leibniz Zentrum für
Archäologie},addressline={Ludwig-Lindenschmit-Forum
1},city={Mainz},country={Germany},countrysep={,},postcode={55116},postcodesep={}}
\affiliation[2]{organization={University of North
Carolina},addressline={104 South Road, 225 Geology
Building},city={Chapel Hill,
NC},country={US},countrysep={,},postcode={27599-3315},postcodesep={}}
\affiliation[3]{organization={CONICET-CADIC},addressline={AV. B.
Houssay,
200.},city={Ushuaia},country={Argentina},countrysep={,},postcode={9410},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This study provides a reassessment of magnesium to calcium (Mg/Ca)
ratios in Atlantic limpet shells to determine past sea surface
temperatures (SST). While \emph{Patella depressa} along the Spanish
shoreline and \emph{Patella caerulea} in the Mediterranean have
repeatedly produced reliable correlations between SST and Mg/Ca ratios,
this relationship is not the case for other patelloid species.
Particularly, \emph{Patella vulgata} but also \emph{Nacella deaurata}
have been studied using Mg/Ca ratios with mixed or contrary results. In
this study, we present elemental maps of these two species as well as
\emph{Nacella magellanica} together with oxygen isotope ratios
(\(\delta\)\textsuperscript{18}O) that confirm a good relationship with
SST. Our dataset also reassesses a specimen which was previously
unsuccessful in providing significant correlations between
\(\delta\)\textsuperscript{18}O values and Mg/Ca ratios. By reassessing
these species and including modern and archaeological specimens (n=12)
from three patelloid species (\emph{P. vulgata}, \emph{N. deaurata}, and
\emph{N. magellanica}) we further add to the growing set of evidence for
the reliable use of Mg/Ca ratios to detect palaeotemperature change. As
a result, these species can in the future serve to determine ontogenetic
age and season of capture as well as to reveal locations of interest
within the growth record (i.e.~annual temperature minima and maxima) for
targeted \(\delta\)\textsuperscript{18}O and clumped isotope analysis.
\end{abstract}

\begin{keyword}
    Sclerochronology \sep Patella \sep Nacella \sep LIBS \sep 
    Mg/Ca
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{Introduction}

Limpet shells are commonly found within archaeological sites and past
shorelines due to their robust carbonate structure and the long-term use
of limpets as a marine food source
\citep[\citet{Robson2023-oz}]{Alvarez2011-rm}. They have been
successfully studied in the context of coastal subsistence economies and
site occupation on seasonal
\citep{Shackleton1973-ij, Parker2020-oy, Bosch2018-ud} or long-term
scales \citep{Ortiz2015-mr}, as well as archives of palaeotemperature
\citep{Fenger2007-gf, Surge2012-ba, Wang2012-ee, Colonese2012-ct, Ferguson2011-zl}.
Determining past sea surface temperature (SST) change recorded in
patelloid limpet shells largely relies on the measurement of
\(\delta\)\textsuperscript{18}O values within the calcitic parts of the
shell, but attempts have been made to also use elemental ratios, such as
magnesium to calcium (Mg/Ca) ratios
\citep{Graniero2017-io, Ferguson2011-zl}, to have an alternative measure
that potentially provides a more accurate SST estimate than
\(\delta\)\textsuperscript{18}O-values,as the latter proxy is affected
by changes in the mixing of fresh- and saltwater. In addition, the data
acquisition of elemental ratios --- either through
laser-ablation-isotope-ratio-mass-spectrometry (LA-ICP-MS) or laser
induced breakdown spectroscopy (LIBS), can be much faster and
cost-effective, increasing the number of specimens that can be studied
overall \citep{Durham2017-fh, Hausmann2023-ih}.

Problematically, links between Mg/Ca ratios and SST changes in many
marine mollusc shells, including limpets, have shown to be unreliable
\citep{Surge2008-ri, Wanamaker2008-zl, Schone2010-yl, Freitas2012-tx, Graniero2017-io, Poulain2015-dg, Vihtakari2017-wd}.
This is particularly the case where there is little available additional
information on metabolic processes, organic components of the shell
matrix, intra-increment and intra-shell variability, and growth rates,
which can independently and unpredictably affect Mg/Ca ratios and
confound their interpretation as a temperature proxy. Confusingly, in
some cases multiple different temperature equations have been found for
the same species (see Freitas \emph{et al}. \citeyearpar{Freitas2012-tx}
and Vihtakari \emph{et al}. \citeyearpar{Vihtakari2017-wd} and
references therein). Coeval specimens sharing one locality can also show
differences in their relation to SST \citep{Hausmann2019-fi}. Where the
use of Mg/Ca ratios as a palaeotemperature proxy was successful,
anomalous patterns in some specimens had to be filtered out by hand,
reducing the overall robustness of the results of those successful
studies \citep{Ferguson2011-zl}.

Recent research particularly of \emph{Patella} spp. in the Mediterranean
and southwest Europe have provided promising results
\citep{Hausmann2019-fi, Garcia-Escarzaga2015-jc, Garcia-Escarzaga2018-nf}.
However, there remains a lack of clarity for Atlantic limpet species,
particularly since it was shown, that the Mg/Ca ratio in their shells
are not reliable recorders of palaeotemperature \citep{Graniero2017-io}.
Using modern reference material Graniero \emph{et al}.
\citeyearpar{Graniero2017-io} tested multiple elemental ratios (Mg/Ca,
Sr/Ca, Li/Ca, Li/Mg, and Sr/Li) acquired via LA-ICP-MS measurements with
no ratios showing correlations with corresponding SST changes nor any
sinusoidal patterns that would indicate a seasonal environmental control
of elemental ratios. This study included two species also used in our
study (\emph{Patella vulgata}, \emph{Nacella deaurata}), with one
\emph{P. vulgata} specimen (i.e.~specimen ORK-LT5) being analysed by
both. We also added specimens of the species \emph{Nacella magellanica,}
which --- similar to the other two species --- is also an established
palaeothermometer using \(\delta\)\textsuperscript{18}O-values
\citep{Colonese2012-ct, gutiérrez-zugasti2017, Nicastro2020-ih}.

In this study, we repeat and expand upon the analysis of Atlantic
limpets to determine the reliability of Mg/Ca ratios as
palaeotemperature proxies. To do this, we sampled a set of previously
published and unpublished limpet specimens dating to modern and
archaeological contexts using LIBS. LIBS allows us to carry out 2D
imaging of entire shell sections, which facilitates navigation of the
complex elemental structure of the shell and a better separation of the
external, temperature-related changes from the internal and less
understood factors that influence the Mg/Ca ratio. By working with
already published datasets, we were able to simultaneously avoid costs
for new high-resolution \(\delta\)\textsuperscript{18}O-data, to use
real-world examples, and to also provide pilot-data for areas of
existing research interest. Generally establishing the usefulness and
reliability of Mg/Ca ratios as SST proxy in limpets provides a platform
for future research and an important steppingstone to a better
understanding of elemental ratios in other marine mollusc shells.

\section{Materials and Methods}\label{Methods}

\subsection{Limpet specimens}\label{limpet-specimens}

The analysed specimens, their origin and respective studies with
research background information can be found in Table
\hyperref[Table_1]{1}. Here we will briefly summarise their contextual
information, which can be accessed in more detail from the respective
studies \citep{Nicastro2020-ih, Surge2012-ba, Graniero2017-io}. While
those studies also included other specimens, their accessibility or
state of preservation did not allow them to be re-analysed.

\phantomsection\label{Table_1}
\fontsize{8pt}{8pt}\selectfont

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1389}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2361}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1250}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{Overview of the modern and archaeological limpet specimens
analysed in this study.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Modern & \citep{Nicastro2020-ih} & \emph{N. deaurata} & Cambaceres Bay &
ND-1016-3 & \(\delta\)\textsuperscript{18}O and
\(\delta\)\textsuperscript{13}C \\
& & & - Tierra del & ND-1016-4 & analysis \\
& & \emph{N. magellanica} & Fuego (AR) & NM-1016-1 & \\
& & & & NM-1016-3 & \\
& & & & & \\
& \citep{Graniero2017-io} & \emph{P. vulgata} & Rack Wick Bay - &
ORK-LT5 & \(\delta\)\textsuperscript{18}O and
\(\delta\)\textsuperscript{13}C \\
& & & Westray (UK) & & analysis; \\
& & & & & Mg, Li, Sr, Ca \\
& & & & & \\
& & & & & \\
Archaeological & \citep{Surge2012-ba} & \emph{P. vulgata} & Rack Wick
Bay - & QG1-7188-1 & \(\delta\)\textsuperscript{18}O and
\(\delta\)\textsuperscript{13}C \\
& & & Westray (UK) & QG1-7189-2 & analysis \\
& & & & QG2-1061-1 & \\
& & & & QG2-1064-1 & \\
& & & & QG2-7180-1 & \\
& & & & QG2-7180-2 & \\
& & & & QG1-7246-1 & \\
\end{longtable}

\normalsize

Specimens were sourced from two main locations: the Beagle Channel in
Tierra del Fuego (Argentina) and the island of Westray in the Orkney
archipelago of Scotland (UK). Four modern specimens were collected from
the Beagle Channel in Cambaceres Bay in October 2016. The area receives
around 570 mm of annual precipitation, but the channel waters are
predominantly influenced by marine currents and in the period of October
2015 to October 2016 stayed around 30.7 ± 0.7 psu (practical salinity
units) \citep{Nicastro2020-ih}. SSTs range from 30.5ºC in August to
31.9ºC in February (Figure~\ref{fig-SSTs} -- red line).

On Westray one modern shell (ORK-LT5) was collected in Rack Wick Bay in
August 2009 and has been previously analysed for
\(\delta\)\textsuperscript{18}O and \(\delta\)\textsuperscript{13}C
values, as well as for elemental ratios via LA-ICP-MS
\citep{Graniero2017-io}. Westray lies about 70 km north of the Scottish
mainland and experiences virtually no freshwater input with salinity
values of 34.0 -- 34.5 psu \citep{Inall2009-ho}. The annual SST ranges
from 34.4ºC in March to 35.3ºC in August (Figure~\ref{fig-SSTs} -- blue
line).

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-SSTs-1.pdf}

}

\caption{\label{fig-SSTs}Monthly sea surface temperature values averaged
from 2007 to 2023. Values were sourced from \citet{Good2020-nl}.}

\end{figure}%

We also studied 7 archaeological specimens of \emph{P. vulgata} from the
Quoygrew site on Westray dating to 900 -- 1200 CE. These shells have
previously been studied to characterise seasonal temperature change
during the Medieval Climate Anomaly (810 -- 1229 CE)
\citep{Surge2012-ba}.

\subsection{Methods}\label{methods}

\subsubsection{Oxygen isotope ratios}\label{oxygen-isotope-ratios}

High-quality \(\delta\)\textsuperscript{18}O values were acquired from
previous publications on \emph{P. vulgata}
\citep{Surge2012-ba, Graniero2017-io}, \emph{N. deaurata} and \emph{N.
magellanica} \citep{Nicastro2020-ih}, in which detailed descriptions of
the methodology for the stable isotope analyses can be found. In short,
all shells were sectioned along the maximum growth axis to expose the
internal growth structure. From the exposed section carbonate powder
samples were acquired using a micromill with resolutions of 100--200 µm
and by targeting the calcitic layers of the shells (m+2 and m+3, see
\citep{Fenger2007-gf}). The exact location of the sampled carbonate
powder is not available apart from the distance to the growth edge,
where reported. Analytical precision of the isotope ratios was
consistently at 0.05--0.10‰ and values are reported in per mil units (‰)
relative to the VPDB (Vienna Pee Dee Belemnite) standard.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_iso-1.png}

}

\caption{\label{fig-Nac_iso}Isotope values of \emph{Nacella} spp. limpet
shells from Cambaceres Bay.}

\end{figure}%

Figure~\ref{fig-Nac_iso} and Figure~\ref{fig-Pat_iso} show the
previously acquired sequential \(\delta\)\textsuperscript{18}O-values of
limpet shells by species (\emph{Nacella} spp., and \emph{Patella
vulgata}, respectively). Due to the high-resolution sampling, all shells
show seasonally resolved and multi-year quasi-sinusoidal records of SST
change, with \emph{Nacella} spp. shells living 2--3 years and \emph{P.
vulgata} living regularly over 3 years. In those instances where longer
lifespans resulted in slow growth rates below 1 mm/yr (i.e.~specimen
QG1-7246-1, Figure~\ref{fig-Pat_iso} b), the temporal-resolution is
reduced from more than a dozen of samples per year, to only around 7.
Together they show a range of lifespans, with some comparatively long
lived for their size (QG1-7246-1, 12 years within 16 mm of growth)
making them particularly susceptible to within-shell time-averaging
because of slow growth rates. That said, the majority yielded
sufficiently fast growth to nearly capture seasonal amplitudes of SST
with ages of 2--3 years (n=3), 3--4 years (n=2), and 5 years (n=2).

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_iso-1.png}

}

\caption{\label{fig-Pat_iso}Isotope values of \emph{Patella vulgata}
limpet shells from Quoygrew (QG, \emph{a-g}) and Rack Wick Bay (ORK,
\emph{h}) on Westray, Orkney.}

\end{figure}%

\subsubsection{Mg/Ca ratios}\label{mgca-ratios}

Mg/Ca elemental imaging was carried out using LIBS at the Leibniz
Zentrum für Archäologie (Mainz-Germany) following previously published
methods \citep{Hausmann2023-ih}. The imaging involves the ablation of
the shell section using an infrared (1064 nm) laser (1---2 mJ, 100 Hz)
onto a surface area of 20---30 µm to generate a plasma plume. This plume
emits light, which is measured using a synchronised spectrometer (delay
time = 0.5 µs, integration time = 1 µs). The resulting light spectrum
quantifies the emission lines of magnesium (MgII: 279.553 nm) and
calcium (CaII: 315.887 nm) to determine their intensity ratio. While
these two peaks alone do not represent the molar concentrations of both
elements (as opposed to calibration free LIBS; e.g.
\citep{Martinez-Minchero2023-pq}), the intensity ratio is linearly
correlated to the molar concentration \citep{Hausmann2017-oa} and can be
used as a reliable indicator of Mg/Ca variation within the shell
carbonate. All values reported here are in arbitrary units based on the
ratio of intensities of the peaks above.

Using this system we carried out elemental imaging of the shell
specimens at a resolution of 50--100 µm distance between sampling spots.
Each spot was irradiated 10 times with the first 3 spectra discarded as
a cleaning step and the remaining spectra summed to get an average Mg/Ca
intensity ratio for each sample spot. Subsequently, we re-sampled the
section using a line scan at 10 µm resolution. This leads to an overlap
between sample spots, but also allows for a continuous record without
gaps. Intensity ratios were filtered for cases with high relative
standard deviation (i.e., more than 10\%). These occurred in places
where the previous sampling procedures for carbonate powder as part of
the oxygen isotope analysis left an uneven sample surface introducing
variability in the plasma generation and thus uncertainty in the data of
these locations.

\subsection{Dynamic Time Warping}\label{dynamic-time-warping}

Because it is not possible to directly associate the exact sample
locations used in the analysis of oxygen isotope ratios with the LIBS
data, we used dynamic time warping (DTW) to align the time series of
\(\delta\)\textsuperscript{18}O values and Mg/Ca ratios. DTW is an
algorithm that measures similarity between two proxy sequences, which
may vary in sampling resolution or interval. By stretching or
compressing sections of the series, DTW finds the probable alignment
between the two sequences. This allows us to compare the proxy datasets
more effectively, ensuring that the temporal dynamics of each shell are
accurately matched despite possible discrepancies in sampling intervals
or rates. We applied the DTW algorithm using the \texttt{dtw} package in
R \citep{Giorgino2009-sj, R_Core_Team2020-mk}, which provides a robust
framework for aligning time series data. This can involve selecting
appropriate distance measures and constraints to ensure meaningful
alignment. We used the default settings for symmetric step patterns and
euclidean distance measures (see code in Hausmann et
al.\citeyearpar{hausmann2024confirmation}).

\section{Results}\label{Results}

\subsection{\texorpdfstring{\emph{Patella
vulgata}}{Patella vulgata}}\label{patella-vulgata}

The elemental imaging of \emph{P. vulgata} shells consistently showed
repeating patterns of Mg/Ca intensity ratio changes with high Mg/Ca
intensity ratios indicating high temperature periods and low ratios
indicating low temperature periods. As examples,
Figure~\ref{fig-Pat_LIBS} shows the 2D-distribution of elemental ratios
across the entire section of ORK-LT5 and the preserved anterior side of
QG1-7188-1 (additional maps can be found in the supplementaries). The
repeating patterns are found across the calcitic layers that are
exterior to the myostracum (layers m+2 and m+3) as well as the layers
that are interior (m--2)(see Fenger et al. \citeyearpar{Fenger2007-gf}
for a detailed explanations of microstructural layers in \emph{P.
vulgata}). The layer m-1, which is also interior, consists of aragonite
and is seen in Figure~\ref{fig-Pat_LIBS} in grey, as its Mg/Ca intensity
ratio consistently falls below the range of interest (0.2 and above).
Compared to other patelloid shells, some of whose interior is almost
entirely made of aragonite, this layer is very thin. \newline

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_LIBS-1.pdf}

}

\caption{\label{fig-Pat_LIBS}Example LIBS maps and line scans of
analysed specimens. (\emph{a+c}) ORK-LT and (\emph{b}+\emph{d})
QG1-7188-1. Note that the ranges of Mg/Ca intensity ratios are not
identical but are chosen arbitrarily to increase contrast. Areas in grey
consist of low-Mg aragonite and fall outside the chosen colour-range.
White dashed lines indicate location of line scans.}

\end{figure}%

Both specimens in Figure~\ref{fig-Pat_LIBS} show increased Mg/Ca
intensity ratios towards the outside of the shell and some degree of
intra-increment variability. This is particularly visible in
Figure~\ref{fig-Pat_LIBS} b, where Mg/Ca intensity ratios range from
\textasciitilde0.5--0.7 in the first year of growth (9-\/--12 mm on the
y-axis). A similar pattern is visible in the interior m-2 layer of
ORK-LT5 (Figure~\ref{fig-Pat_LIBS} a), which increase towards the
anterior of the shell (0--10 mm on the x-axis).

Line scans also indicate well the quasi-sinusoidal change of Mg/Ca
intensity ratios expected based on the stable isotope data and elemental
imaging. That said, some of the variability of Mg/Ca intensity ratios
within one season such as the summer period of QG1-7188-1 between 0.5
and 2.0 mm distance to the shell edge (Figure~\ref{fig-Pat_LIBS} d), is
visible in a more dramatic manner than it appears in the 2D image. This
shows the downside of line-scanning as opposed to the broader milling
approach used for the analysis of carbonate powder
\citep{Ferguson2011-zl}.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_Comp-1.png}

}

\caption{\label{fig-Pat_Comp}Comparison of LIBS line-scan and sequential
\(\delta\)\textsuperscript{18}O-values of specimen ORK-LT5}

\end{figure}%

Figure~\ref{fig-Pat_Comp} shows the line scan of ORK-LT5 in comparison
to its \(\delta\)\textsuperscript{18}O values (similar graphs for other
specimen can be found in the supplementaries). The measurements of the
distance to the shell edge are not entirely identical, most likely
because one growth increment can have a range of distances to the edge,
depending on where one measures, and a line scan along the interior
would be shorter than a line scan along the exterior. That said,
increases and decreases of both proxies seem to mirror each other and to
be well aligned.

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Pat_Corr-1.png}

}

\caption{\label{fig-Pat_Corr}Correlation graphs for Mg/Ca and
\(\delta\)\textsuperscript{18}O values in \emph{P. vulgata} specimens.}

\end{figure}%

Aligning the records programmatically using dynamic time warping,
allowed us to compare both proxies for each specimen directly to
determine specimen-specific equations and quantify the correlation of
both proxies (Figure~\ref{fig-Pat_Corr}). The equations for all
specimens are different but the majority (barring one: QG1-7246-1) do
seem to cluster around shared parameters
(\(\delta\)\textsuperscript{18}Ο = 0.6 -- 0.065 * Mg/Ca). The various
coefficients of determination (R\textsuperscript{2}) range between 0.81
(QG1-7246-1) and 0.95 (QG1-7189-2). These specimens are also the oldest
and youngest specimens, respectively, suggesting that lower growth rates
and time averaging had a negative effect on the correlation of both
proxies. The mean R\textsuperscript{2} value of the individual \emph{P.
vulgata} specimens is 0.89, suggesting a generally good fit between the
two proxies. P-values are overwhelmingly significant and consistently
below p\textless1e\textsuperscript{--6}.

\subsection{\texorpdfstring{\emph{Nacella}
spp.}{Nacella spp.}}\label{nacella-spp.}

The LIBS data for \emph{Nacella} spp. specimens is less straightforward
with shells being much thinner than \emph{P. vulgata} and thus patterns
being more difficult to discern. Both specimens (ND-1016-3 and
NM-1016-3) shown as examples in Figure~\ref{fig-Nac_LIBS} (see
additional graphs in supplementaries) also experience increases of Mg/Ca
intensity ratio towards the exterior of the shell, similar to the
\emph{P. vulgata} specimens from Orkney. This intra-increment
heterogeneity combined with the thinness of the shell, further
complicates their analysis compared to \emph{P. vulgata} or other
\emph{Patella} species \citep{Hausmann2019-fi}. Nevertheless, repeating
patterns were visible in the anterior section of the shell, which were
also mirrored in the inner layers (m-2) of specimens NM-1016-3
(Figure~\ref{fig-Nac_LIBS} b).

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_LIBS-1.pdf}

}

\caption{\label{fig-Nac_LIBS}Example LIBS maps and line scans of
analysed \emph{Nacella} spp. specimens. (\emph{a+c}) ND-1016-3 and
(\emph{b}+\emph{d}) NM-1016-3. Note that the ranges of Mg/Ca intensity
ratios are not identical but are chosen arbitrarily to increase
contrast. Areas in grey consist of low-Mg aragonite and fall outside the
chosen colour-range. Note the inset in (\emph{b}) where the shell edge
has been resampled using a higher resolution (50 µm instead of 100 µm)
to better understand the Mg/Ca ratios at the shell edge. White dashed
lines indicate location of line scans.}

\end{figure}%

Interestingly, the Mg/Ca intensity ratios in the \emph{Nacella} spp.
shells were much lower (0.05--0.30) than those seen in other patelloid
species (e.g.~0.3--0.7 in \emph{P. vulgata} above, or 0.5--1.5 in
Hausmann \emph{et al}. \citeyearpar{Hausmann2023-ih}) using the same
emission lines. While the intensity ratio depends chiefly on the chosen
emission lines to calculate the proportion of the chosen magnesium to
the chosen calcium peak, the low ratios still indicate lower
concentrations than seen elsewhere or seen only in aragonitic parts of
e.g.~\emph{Patella} shells. Similar observations have been made before
\citep{Graniero2017-io}. Strong intra-increment heterogeneities are
visible along the exterior of the shells (m+3 layer), where Mg/Ca
intensity ratios are higher, particularly in specimen ND-1016-3
(Figure~\ref{fig-Nac_LIBS}). The step from lower to higher along the
increment is not gradual, as is the case in NM-1016-3 (see inset of
Figure~\ref{fig-Nac_LIBS} b), but rather stepwise. Sampling across this
boundary in a linear scan would most likely lead to erroneous results.
Also sampling across gradual changes as in NM-1016-3 can influence the
overall correlation graph for a specimen.

Line scans reflect well the changes previously indicated by
\(\delta\)\textsuperscript{18}O values and also indicate about two years
of recorded growth. That said, the 2D imaging suggests that in both
specimens a third summer can be added to the total record, which is not
captured by the line scans.

Figure~\ref{fig-Nac_Comp} shows the line scans of ND-1016-3 and
NM-1016-3 in comparison to their \(\delta\)\textsuperscript{18}O values
(similar graphs for the other \emph{Nacella} spp. specimens can be found
in the supplementaries). Compared to the \emph{P. vulgata} records,
these mirror each other much more clearly, due to the brief period
recorded in the shells and thus a lower chance of misaligning annual
extremes.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_Comp-1.png}

}

\caption{\label{fig-Nac_Comp}Comparison of LIBS line-scan and sequential
\(\delta\)\textsuperscript{18}O-values of \emph{Nacella} spp.
specimens.}

\end{figure}%

The \emph{Nacella} spp. specimens showed similarly clustered
correlations, which interestingly were not grouped by species. In fact,
specimens ND-1016-3 (\emph{Nacella deaurata)} and NM-1016-3 (\emph{N.
magellanica}) are almost identical in their fitted equations
(Figure~\ref{fig-Nac_Corr}). Both also share a high coefficient of
determination (R\textsuperscript{2}) of 0.95 and 0.92 respectively, with
the remaining specimens being somewhat lower (R\textsuperscript{2} =
0.76 for ND-1016-4 and R\textsuperscript{2}=0.84 for NM-1016-3). The
mean R\textsuperscript{2} value for each \emph{N.} species is 0.86 for
\emph{N. deaurata} and 0.88 for \emph{N. magellanica} with p-values for
all speciemns being far below 0.0001. Why two specimens of different
species are so well aligned is not entirely certain and randomness
cannot for now be ruled out. That said, since they both have a high
R\textsuperscript{2} value, it might just be that they were both
minimally affected by factors other than SST, including the sampling
location of our line scans.

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-Nac_Corr-1.png}

}

\caption{\label{fig-Nac_Corr}Correlation graphs for Mg/Ca and
\(\delta\)\textsuperscript{18}O values in \emph{Nacella} spp. specimens}

\end{figure}%

\section{Discussion}\label{Discussion}

The Mg/Ca ratios from the 2D elemental imaging present results
suggesting that SST is the main factor influencing variations in Mg/Ca
ratios in three Atlantic limpet species. Annually repeating patterns of
increasing (spring/summer) and decreasing (autum/winter) SST were also
visible in the line scans, albeit less straightforward than in the
2D-data. Nevertheless, the resulting mean R\textsuperscript{2}-values
comparing Mg/Ca and δ\textsuperscript{18}O values for each species were
high (\emph{P. vulgata R\textsuperscript{2}}=0.89, \emph{N. deaurata
R\textsuperscript{2}}=0.86\emph{, N. magellanica
R\textsuperscript{2}}=0.88). These results are similar to previously
published data on other \emph{Patella} species shells (Table
\hyperref[Table_2]{2}): R\textsuperscript{2} values of \emph{P.
depressa} lie between 0.78 and 0.87 \citep{Garcia-Escarzaga2021-ij}, for
\emph{P. caerulea} from the Mediterranean between 0.78 and 0.96 ---
barring one outlier of 0.33 (specimen MP64A), which had increased Mg/Ca
ratios in increments with rich organic contents \citep{Hausmann2019-fi}.

\phantomsection\label{Table_2}
\fontsize{8pt}{8pt}\selectfont

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1471}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1618}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1324}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1029}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2059}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2500}}@{}}
\caption{Correlations of Mg/Ca ratios with
\(\delta\)\textsuperscript{18}O values found in other studies. Bold
R\textsuperscript{2} values are interpreted as outliers. * indicates
samples that were used to determine a shared R\textsuperscript{2}-value
of 0.79.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Locality
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Specimen
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
R\textsuperscript{2}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Locality
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Specimen
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
R\textsuperscript{2}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\emph{P. depressa} & Northern Spain & LAN541 & 0.87 &
\citep{Garcia-Escarzaga2021-ij} & \\
& & LAN545 & 0.86 & & \\
& & LAN554 & 0.78 & & \\
& & LAN559 & 0.82 & & \\
\emph{P. caerulea} & Croatia & ISTPC1 & 0.9 & \citep{Hausmann2019-fi}
& \\
& & ISTPC2 & 0.84 & & \\
& Crete & AF1911A & 0.91 & & \\
& & AF3003A & 0.92 & & \\
& Israel & AKKPC2 & 0.96 & & \\
& & AKKPC3 & 0.89 & & \\
& & FRMPC1 & 0.84 & & \\
& & FRMPC2 & 0.96 & & \\
& Libya & MO31A & 0.83 & & \\
& & \textbf{MP64A} & \textbf{0.33} & & \\
& & MP67A & 0.96 & & \\
& & MP68A & 0.81 & & \\
& Malta & MA10 & 0.82 & & \\
& Tunisia & TUNPC1 & 0.81 & & \\
& & TUNPC2 & 0.78 & & \\
& Turkey & ANTPC1 & 0.95 & & \\
& & ANTPC2 & 0.93 & & \\
& & KIZPC1 & 0.94 & & \\
& & KIZPC2 & 0.86 & & \\
\emph{P. rustica} & Gibraltar & \textbf{JL1} & \textbf{0.02} &
\citep{Ferguson2011-zl} & \\
& & JL2 & 0.8* & & \\
\emph{P. caerulea} & Gibraltar & JM00 & 0.69* & & \\
& & JM30 & 0.83* & & \\
\emph{P. vulgata} & Orkney & ORK-LT5 & not reported, here 0.88 &
\citep{Graniero2017-io} and this study & \\
& & QG1-7188-1 & 0.88 & this study & \\
& & QG1-7189-2 & 0.95 & & \\
& & QG1-7246-1 & 0.81 & & \\
& & QG2-1061-1 & 0.90 & & \\
& & QG2-1064-1 & 0.90 & & \\
& & QG2-7180-1 & 0.87 & & \\
& & QG2-7180-2 & 0.90 & & \\
\emph{N. deaurata} & Tierra del Fuego & ND-1016-3 & 0.95 & & \\
& & ND-1016-4 & 0.76 & & \\
\emph{N. magellanica} & Tierra del Fuego & NM-1016-1 & 0.92 & & \\
& & NM-1016-3 & 0.84 & & \\
\end{longtable}

\normalsize

\emph{P. rustica} has two studied specimens (JL1 and JL2) with
R\textsuperscript{2}-values of 0.02 and 0.80, respectively
\citep{Ferguson2011-zl} (see also Supplementaries). Together with the
data from two \emph{P. caerulea} shells (JM00 and JM30 with
R\textsuperscript{2}-values of 0.69 and 0.83), the data from JL2 was
used to determine a general Mg/Ca-SST relationship with an
R\textsuperscript{2} value of 0.79. JL1 was disregarded in that
calculation, as well as the second year of growth in JM00, which
produced anomalous data. While no additional information about those
specimens is available, it seems likely that the discarded Mg/Ca ratios
from JM00, which were interpreted as an ontogenetic trend, were
potentially influenced by the same intra-increment heterogeneity we see
in some of our shell specimens as well. Since these shells were analysed
by homogenising the carbonate powder of the m+2 layer and the more
exterior m+3 layer, these heterogeneities could have potentially
increased the Mg/Ca ratio unpredictably, leading to what might look like
an ontogenetic trend in a linear dataset. Whether these anomalies
affected JL1 in the same way, albeit stronger, is not clear but it might
well be possible. Previous studies of \emph{P. caerulea} found similar
patterns of anomalous Mg/Ca values in the m+3 layer
\citep{Hausmann2019-fi}. These areas can be avoided using a
2-dimensional approach and if the growth increment is sufficiently wide.
These heterogeneities within growth increments are little understood but
some research points towards a zonation of the mollusc's mantle and
along the growth edge \citep{Lazareth2013-jf}. This zonation could
differentially influence the transport of Mg ions to the extrapallial
fluid and into the shell.

Another shell whose results we aim to explain further is ORK-LT5 (see
Supplementary Information), which has previously been shown to have no
substantial correlation with \(\delta\)\textsuperscript{18}O values or
SST by implication \citep{Graniero2017-io}. In contrast to other
specimens in this study, we were able to resample this specimen and
acquire Mg/Ca data on a slightly higher resolution (20--30 µm spots in a
100 µm raster compared to 50 µm spots spaced at 150--300 µm) and using 2
dimensions, as well as using a continuous high-resolution line-scan at
(20--30 µm spots spaced at 10 µm) (Figure~\ref{fig-ORK_sub}). These
additional data revealed previously missed summer peaks at around 2--3
mm distance to the shell edge (grey box in Figure~\ref{fig-ORK_sub}) and
revealed additional growth at the very edge (dashed box in
Figure~\ref{fig-ORK_sub}) mirroring the change towards warmer
temperatures at the time of collection (August). Including these missing
parts lead to a better alignment of both records and a resulting
R\textsuperscript{2}-value of 0.88 (Figure~\ref{fig-ORK_sub}).

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-ORK_sub-1.pdf}

}

\caption{\label{fig-ORK_sub}Comparison of previous and new Mg/Ca data
for ORK-LT5 explaining the previously unsuccessful correlation of Mg/Ca
and \(\delta\)\textsuperscript{18}O-values.}

\end{figure}%

We were not able to resample the \emph{Nacella} spp. specimens in
Graniero \emph{et al.} \citeyearpar{Graniero2017-io}, which had
similarly unrelated records of \(\delta\)\textsuperscript{18}O values
and Mg/Ca ratios and pointed to no relationship between elemental ratios
and SST at all. However, we were able to source other modern specimens
of the same species from Tierra del Fuego, which shall serve here as an
indicator. Similar to other patelloid shells, the results of the
\emph{Nacella} specimens in our study were successful with high
R\textsuperscript{2} values (0.76--0.95) and suggest Mg/Ca as a reliable
proxy for SST (Figure~\ref{fig-Nac_Comp} and supplementaries). Why these
results are different from the studied \emph{Nacella} spp. specimens in
Graniero \emph{et al}. \citeyearpar{Graniero2017-io} is not obvious, but
the occasional alignment of the Mg/Ca and
\(\delta\)\textsuperscript{18}O records of their shells (Figure 6 in
Graniero \emph{et al}. \citeyearpar{Graniero2017-io}) indicate that
there is some relationship that --- potentially due to the low sampling
resolution or intra-increment heterogeneity --- is not perfectly
visible, leading to an overall result of an uncorrelated relationship
between Mg/Ca ratios and SST.

\subsection{Temperature correlation}\label{temperature-correlation}

We further related the Mg/Ca ratios of the modern shells to actual SST
data from the respective localities (Figure~\ref{fig-SSTs}). The data
for ORK-LT5 from Orkney were cut to the time period between the start of
our available SST data (2007) and the time of collection (August 2009).
For the \emph{Nacella} spp. from Cambaceres Bay we chose an overlapping
time period, to make the results more comparable between the specimens,
which is the lifespan of the youngest specimen NM-1016-3 from January
2015 to October 2016, the time of collection for all 4 specimens.

As indicated by the stable isotope values, no reliable shared
relationship between the data can be established, but each shell's Mg/Ca
data individually have a good correlation with SST (R\textsuperscript{2}
values between 0.93 to 0.96 with p-values consistently below 0.001)
(Figure~\ref{fig-SST_Corr}), which are even better than the correlation
between elemental ratios and stable isotope values. This improvement is
most likely due to the simple and sinusoidal SST curve, which is easier
to fit to other data.

As a result of the different correlations between Mg/Ca ratios and SST
in the modern shells, we did not attempt to use Mg/Ca as a direct proxy
for absolute SST values in ºC for the archaeological shells from
Quoygrew, but rather as a proxy for relative SST change
(i.e.~warmer/colder). In future analyses, these high resolution and
rapidly acquired records of relative SST can guide the sampling for
oxygen isotope or clumped isotope analysis and reduce analytical costs
per shell, thus increasing the quantity of shells that can be analysed
per study.

\begin{figure}[H]

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-SST_Corr-1.png}

}

\caption{\label{fig-SST_Corr}Comparison of Mg/Ca and SST in all modern
shells.}

\end{figure}%

\section{Conclusion}\label{conclusion}

This study uses LIBS as a novel analytical tool to present
high-resolution and 2D information on Mg/Ca concentrations in modern and
archaeological patelloid shells and their relationship to sea surface
temperature. The results for \emph{P. vulgata}, \emph{N. deaurata}, and
\emph{N. magellanica} were consistently promising with mean
R\textsuperscript{2} values of 0.89, 0.86, and 0.88 respectively.

LIBS elemental imaging provided insight into elemental anomalies found
particulary in the external parts (m+3 layer) of the shell, which
introduce intra-incremental heterogeneity and which cannot be reconciled
with modern SST changes or SST changes recorded in the
\(\delta\)\textsuperscript{18}O values of the same shells. Our 2D
approach allowed us to avoid these areas of intra-increment
heterogeneity and focus on areas that are predominantly controlled by
changes in SST.

These results are particularly promising for future analyses of
patelloid shells for the study of season of collection in archaeological
sites as well as the reconstruction of sub-annual climatic conditions in
the past. Shells of the species studied here are widely available along
the European Atlantic façade and the Argentinian coastline and make up a
significant part of marine mollusc remains in archaeological sites of
those regions
\citep{Colonese2011-ab, Villagran2011-ld, Zangrando2016-yl}. By using
LIBS as a screening tool for targeted \(\delta\)\textsuperscript{18}O
analysis, focusing on annual minima and maxima revealed in the Mg/Ca
record, researchers can improve cost and time efficiency. This novel
method will enable them to increase sample numbers and produce larger,
more in-depth climatic studies than previously possible.

\section{Acknowledgements}\label{acknowledgements}

Niklas Hausmann was supported by the German Research Foundation under
the Emmy Noether Program (project number: 439799406). We thank the
Applied Photonics Lab of the Institute of Elec- tronic Structure and
Laser of the Foundation for Research and Technology - Hellas for close
support with the LIBS system. We are also grateful for two anonymous
reviewers who helped to improve the manuscript.

  \bibliography{bibliography.bib}

\end{document}